---
title: "ETC 3250 Lab 2 2017 - Code"
author: "Mojdeh Shirazi-Manesh"
date: "2 August 2017"
output: pdf_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  error = FALSE, 
  collapse = TRUE,
  comment = "#",
  fig.height = 4,
  fig.width = 8,
  fig.align = "center",
  cache = FALSE
)
```

```{r }
library(readr)
library(plyr)
library(dplyr)
library(tidyr) #
library(knitr) #
library(ggplot2) # for graphics
library(gridExtra)

dataset <- read.csv("speed-dating-data.csv")
dim(dataset)

DT<-select(dataset,wave, iid, id, gender, idg,match, samerace, age_o,race_o, 
           field_cd, race, imprace, imprelig, goal, date, go_out, attr1_1,
           sinc1_1,intel1_1, fun1_1, amb1_1, shar1_1)

glimpse(DT)

summary(DT)

```
### How many people have participated to the speed dating experiment?

```{r }
length(unique(DT$iid))

```

###How many dates each person has participated to? Compute a summary of these numbers


```{r}
dates<-dataset%>%
  select(wave,iid,order)%>%
  group_by(wave,iid)%>%
  tally(order)

dates


dates.summary <- dataset %>%
select(wave,iid,order) %>%
group_by(wave,iid) %>%
summarise(m=mean(order,na.rm=TRUE), s=sd(order,na.rm=TRUE))

dates.summary
```
###Confirm the number of males and females in each wave given in the documentation is correct

. To compute some statistic for each group individually, rather than for the data set as a whole, we can use **aggregate** function from library **dplyr**

. **aggregate(y~x,data,function)**

. function(x) length(unique(x)) : defining new function that comes from R

. y ~ x : y is numeric data to be split into groups according to x variable

```{r}
aggregate(id ~ gender + wave , DT, function(x) length(unique(x)))
```

#Recode Variable
```{r}
# Method 1 : Recode Variable 'Gender'
DT$gender[which(DT$gender == 0)] <- "Female"
DT$gender[which(DT$gender == 1)] <- "Male"
DT$gender <- as.factor(DT$gender)
# Method 2 : Recode Variable 'Match'
DT$match <- as.factor(DT$match)
DT$match <- revalue(DT$match, c("0" = "No", "1" = "Yes"))
```

#ggplot2 for visualization

```{r}

## Frequency of Going Out, Gender, Race
p1 <- ggplot(data=subset(DT,as.character(DT$match)=="Yes"),
             aes(x=factor(go_out),fill=gender)) +
  geom_bar(stat="count",position = position_dodge()) 

p2 <- ggplot(data=subset(DT,as.character(DT$match)=="Yes"),
             aes(x=factor(go_out),fill=gender)) +
  geom_bar(stat="count",position = position_dodge())  +
  facet_wrap(~ race)

grid.arrange(p1, p2, nrow = 2, ncol = 1)

```
